---
title: "Gun Control"
author: "Yuelin Long"
date: "November 16, 2018"
output: github_document
---


```{r include=FALSE}
#Load packages
library(tidyverse)
library(rvest)
library(httr)
```


To quantify how strict are gun controls in different states, we are using:

+ The number of background checks done by U.S. firearms licensees to approximate number of people applied for license; 
+ The number of federal firearms licensees in the U.S. to approximate the number of people get license approved;
+ The proportion of licensees and background checks to indicate approval percentage.

```{r echo=FALSE}

clean_fun = function(address, area){
  readxl::read_xlsx(address, sheet = "Data", range = area) %>% 
  rename(state = X__1) %>% 
  janitor::clean_names()
}


total_pop = clean_fun("./data/population-in-the-states-of-the-us-as-of-2017.xlsx", "B5:C56")

approved_lic = clean_fun("./data/number-of-federal-firearms-licensees-in-the-us-in-2017-by-state.xlsx", "B5:C57") %>% 
  filter(state != "Other Territories")

back_check = clean_fun("./data/nics-background-checks-done-by-us-firearms-licensees-2017-by-state.xlsx", "B5:C56")

approval_rate = inner_join(approved_lic, back_check, by = "state") %>% 
  inner_join(total_pop, by = "state") %>% 
  mutate(app_perc = 100 * number_of_federal_firearms_licensees / number_of_background_checks)

skimr::skim(approval_rate) %>%
            select(variable:stat,value) %>%
            filter(stat != "hist") %>%
            spread(key = stat,value = value) %>% 
  knitr::kable()
```



```{r echo=FALSE}
library(usmap)

plot_usmap(data = approval_rate, values = "app_perc") + 
  scale_fill_continuous(low = "white", high = "red", name = "Approval Percentage (2017)", label = scales::comma) + 
  theme(legend.position = "right") +
  labs(title = "Approval Rate Across different States")

```


```{r}
# website URL
url = "https://lawcenter.giffords.org/scorecard/#rankings"
gun_climate_url = read_html(url)
# extract table and clean data
gun_climate_data = 
	gun_climate_url %>% 
  html_nodes(css = "table") %>% 
  .[[1]] %>% 
  html_table(header = TRUE) %>% 
  as.tibble() %>% 
	janitor::clean_names() %>% 
	rename(
		law_strength = gun_law_strength_ranked, 
		grade_2017 = x2017grade, 
		death_rate_rank = gun_death_rate_ranked, 
		death_rate = gun_death_rate_per_100k) %>% 
	mutate(
    grade_2017 = factor(grade_2017, 
                      levels = c("A", "A-", "B+", "B", "C+", "C", "C-", "D", "D-", "F")), 
    grade_2017 = fct_collapse(grade_2017, 
    													A = c("A", "A-"), 
    													B = c("B+", "B"), 
    													C = c("C+", "C", "C-"), 
    													D = c("D", "D-")), 
    state_abb = state.abb)

```

Using gun legislation strength:

```{r echo=FALSE}
gun_control = gun_climate_data %>% 
  inner_join(approval_rate, by = "state")

ggplot(gun_control, aes(x = app_perc, y = law_strength, color = law_strength)) +
  geom_point() +  
  labs(
    title = "Approval Rate vs Law Strength",
    x = "Percentage of people approved",
    y = "Law Strength"
  )

```

Weird thing happened: legislation strength is not closely related to approval rate as we thought before. 

Then I thought, maybe less people would apply for gun license if the state has strict laws. So I draw a plot between law strength and appoval percentage(how many people get approved license among total population): 

```{r echo=FALSE}
ggplot(gun_control, aes(x = number_of_federal_firearms_licensees/number_of_residents_in_millions, y = law_strength, color = law_strength)) +
  geom_point(aes(size = death_rate)) +  
  labs(
    title = "Approval percentage vs Law Strength",
    x = "Percentage of people approved among total population",
    y = "Law Strength"
  )
```
It works. 

Then I moved on to application percentage (proportion of people applying for the license in total population)

```{r echo=FALSE}
ggplot(gun_control, aes(x = number_of_background_checks/number_of_residents_in_millions, y = law_strength, color = law_strength)) +
  geom_point() +  
  labs(
    title = "Approval Rate vs Law Strength",
    x = "Percentage of people processed among total population",
    y = "Law Strength"
  )
```
Clearly people in different state share equal willing to apply for guns. My first assumption is not correct.

Therefore, the problem should be approval percentage vs. processed percentage:
```{r echo=FALSE}
ggplot(approval_rate, aes(x = number_of_federal_firearms_licensees/number_of_residents_in_millions, y = number_of_background_checks/number_of_residents_in_millions, color = app_perc)) +
  geom_point() +  
  labs(
    title = "Application percentage vs. Proccessed Percentage",
    x = "Percentage of people approved among total population",
    y = "Percentage of people processed among total population"
  )
```

Therefore, I recommend us using Approval percentage vs Law Strength (the second scatterplot). 







